Our Title

### 2015-05-01 draft.tex
Spell checked.

Where appropriate, expected countermeasures, such as shielding of radioactive sources, are discussed within the context of discerning a weak signal in a strong background. In circumstances where it is deemed essential for personnel to monitor the shipment of regular commerce or travel, including but not limited to commercial or private traffic by land, sea, or air, portal detectors and active interrogation is considered. Such locations are qualified for observation which are considered high risk or of strategic value in fielding a conventional response to attack from ISIL on the U.S., allies, or any sovereign territory for which instability is a matter of U.S. National Security. It is assumed that ISIL will prioritize countervalue over counterforce targets, since countervalue targeting is more disruptive and more effective in a campaign of terror.

@@

    Throughout the many conflicts in Iraq in the remainder of the 2000s, AQI underwent many changes in leadership, ideology, and name. These changes, primarily the deaths of much of its upper leadership (including al-Zarqawi), were caused by a combination of military pressure from U.S. troops, multiple disagreements  with bin Laden's al-Qaeda, and backlash from other Sunnis in Iraq \cite{Zelin2014,Kahl2008}. As a result of this, AQI was reorganized as the Islamic State of Iraq (ISI) in 2007, with Abu Bakr al-Baghdadi as it newly appointed leader \cite{Zelin2014,Shadid2010}. Under the new leadership of al-Baghdadi, ISI broke a period of regional detente with al-Qaeda (following the death of Osama bin Laden in August 2011), driving a further wedge between the two groups by al-Baghdadi's refusal to pledge allegiance to the new leader of al-Qaeda, Ayman al-Zawahiri \cite{Al-Jawlani}. In April 2013, this division resulted in  al-Baghdadi announcing that ISI would be extending its reach into Syria, and changing its name to the Islamic State of Iraq and al-Sham, or the Islamic State of Iraq and the Levant \cite{Zelin2014,Al-Hussaini2013}. In June 2014, al-Baghdadi declared ISIL a global Caliphate, with himself as its Caliph \cite{Mortada2014,TheWeek2014}. To understand the significance of this declaration, let us consider the history and role of Caliphs and Caliphates in Islam.

@@

ISIL has many enemies, indeed many more than are graphically depicted in \autoref{fig:ISILRelationships}.  As of January 2015, the anti-ISIL coalition led by the U.S. included 62 nations and groups including the U.S., Canada, Iraq, Jordan, Bahrain, Saudi Arabia, UAE, France, Germany, United Kingdom, Australia, Belgium, Denmark, Italy, Czech Republic, Albania, Netherlands, Estonia, Hungary, Turkey, and Lebanon, all of whom have carried out or supported military operations against ISIL \cite{Wordsworth2015}.  Additionally, although not allied with the U.S. coalition, Syria, Iran, and various Kurdish, Shia, and Sunni militias in Syria and Iraq have opposed ISIL \cite{Mooney2014}.  Despite this broad coalition, there are many unlikely allies in the list which present potential seams for ISIL exploitation  through propaganda or military action.  For example, strikes against ISIL in Syria benefit the Assad regime while U.S. trained rebels could turn on and attack the Assad regime \cite{Shinkman2015}.  Officially, the U.S. supports the removal of the Assad regime albeit without direct military intervention on part of the U.S. \cite{Shinkman2015}.  Similarly, U.S. airstrikes and intelligence have supported Iran and Iraqi Sunni and Shia militias (including several that once fought U.S. forces) in their campaigns against ISIL \cite{Chulov2014}.  While these uneasy alliances have placed pressure on ISIL, it is unclear if they can be sustained over the timeframe that would be necessary to defeat ISIL and prevent their acquisition of a nuclear weapon. Sustaining these alliances would be especially difficult if ISIL were to turn its formidable propaganda machine to breaking some of these fragile links.

@@

The minimum amount of special nuclear material needed to produce a fission weapon is called the critical mass. This value can depend on many different features including isotope, geometry of the material, microstructure of the material, and device design. Given the many variables in play it would be infeasible to include the critical mass of every possible weapons configuration here. However, to give a general idea, \autoref{tab:critical_mass} presents the critical mass of two different designs for several types of weapons. Additionally the yields of fission devices can vary based on many factors including, type of material, device design, amount, efficiency, and many other factors. Yields are typically on the order of kilotons \cite{Moody2014}.

@@

The most important step in the process, from a proliferation standpoint, involves enriching the uranium in 235U. Natural uranium is typically 0.72\% \ce{^{235}U} \cite{Benedict1981}.  If uranium is destined for a power plant, it is typically enriched to 3-5\% \ce{^{235}U}. If on the other hand it is going to be used in a weapon, the uranium must be enriched to greater than approximately 90\% \ce{^{235}U} \cite{Moody2014}.

@@

It should be noted that Syria may have had a reactor for the production of plutonium in the past. However, that site was bombed and destroyed by an Israeli airstrike in 2007 \cite{WorldNuclearAssociation2015a}. It is not impossible that ISIL could obtain the plans for that design. That being said, construction on a nuclear power plant typically takes at least 4 years \cite{NuclearEnergyInstitute2015}.  Furthermore, nuclear reactors are incredibly difficult to hide, as seen by Israel's 2007 air raid against Syria \cite{WorldNuclearAssociation2015a}.  Thus, it seems unlikely that ISIL would be able to complete such a reactor. Therefore, there is a very low probability that ISIL could covertly produce it's own plutonium, and thus little need to monitor for reprocessing signals from ISIL territory.

@@

From the U.S. perspective, 9/11 serves as a case study to understand the shortcomings of IC, subsequent recommended changes, and possible shortcomings facing the IC with respect to ISIL. The 9/11 Report found that there are cultural and organizational factors that hinder strategic communication between U.S intelligence agencies (See \autoref{app:partners} for a disucussion on strategic communication on the international level, which also applies at the interagency level).  Prior to 9/11, U.S intelligence agencies operated in a culture dominated by \enquote{need-to-know}, which implied that organizations were given discretion in sharing data \cite{McConnell2008}. In the context of a dynamic ISIL threat, this means that agencies often have static analyses formed from an incomplete picture and lacking a competing analysis derived from the same data. The divisions between individual agencies render them incapable of accessing, prioritizing, translating, analyzing, and relaying the information for the policymakers to respond to in a timely manner. As a result, information flow is hindered by each agency's networks and databases lacking protocols to interact with other agencies' networks and databases.  Organizational inertia, standard operating procedures, and guidelines hinder the ability to change or implement new ideas. For example, the CIA's culture of risk aversion in the 1990s was detrimental to its intelligence capabilities and prevented them from properly assessing and taking action against the rising threat of terrorism \cite{Zegart2005}.

@@

Several international organizations and treaties are in place to facilite communication flow and prevent the spread of nuclear weapons and technology (see \autoref{app:partners}).  However, communication in these organizations organization can be a challenge, and most of the current infrastructure is designed towards state actors, not the current ISIL threat.  Although not a specific alliance against ISIL or nuclear terrorism, the five eyes agreement (see \autoref{app:intel}) can provide valuable intelligence on possible nuclear activities that ISIL may engage. This group is stable, efficient, and more nimble in intelligence collaboration than the existing international non-proliferation structure.

@@

The number of countries in the coalition is growing with Obama's campaign, but the bar for inclusion is fairly low. While many countries have committed military capabilities to combatting ISIL, some countries join the coalition by merely declaring their intent to contribute. All these member countries don't share the same understanding of whom the real enemy is - ISIL or the Syrian regime led by President Bashar al-Assad. For example, Turkey has urged the coalition to topple the Assad regime as a condition to provide its air bases to carry out attacks against ISIL.

@@

Iran and U.S. shared a common enemy factor in terms of ISIL. As the nuclear talks between the two countries are being hashed out, both countries have committed on fighting ISIL together. This can be a good opportunity for U.S. to combat ISIL more effectively as U.S. cannot disregard Iran's influence and role in Iraq's security landscape and their fight against ISIL.
Meanwhile, collaborating with Israel is considered a gray area and not clear-cut as it seems. Israel has been known to voice their support of ISIL, keeping in line with ISIL'ss common enemy with Israel i.e. Iran. As U.S. and Israel have been a strong ally in the Middle East region, this fact presented a conundrum to the U.S..

@@

If nuclear material is intercepted before detonation, an analysis of the enrichment and contaminants may be achieved at the National Labs (Oak Ridge, Los Alamos). On the scale of weeks to months, it is possible to discern the age of the material post fabrication by analyzing the microstructure, activity, and isotopic composition. Immediately following detonation, the timescale of analysis relies on how quickly radiation monitoring equipment may be delivered to the blast site and how quickly samples are gathered and shipped back to one of 16 radiological facilities equipped to perform nuclear forensics \cite{1446400}. Airplanes capable of sampling large volumes of the dust and gas (Xenon) in the plume generated by the detonation must be performed within 1-2 days of the detonation and may indicate the composition, efficiency, and sophistication of the device. Analysis of materials taken from the ground near the detonation point will take 1-2 weeks to analyze and will confirm and refine the findings of atmospheric monitoring. The identification of the origin of the weapon will depend largely on what is presently known about the yield, sophistication, and efficiency of weapons in global stockpiles. What is exactly known of the capabilities of weapons designed outside of the United States is classified \cite{Glasstone_1964}. Given quick access to the nuclear blast site and a somewhat thorough understanding of the nuclear capabilities of other sovereign powers, the technical analysis of the origin and design of a nuclear blast would be available within 3-10 days, depending on the availability of data. Ongoing research is aimed at developing LIBS technologies which may reduce the nuclear forensics timeline with rapid (\textless 15 hours) debris dissolutions techniques, or techniques which do not require the dissolution of collected debris \cite{Condron}.

@@

While the other weapon models have a fair distance of detection for a short time, the data does not take into account, additional shielding. To prevent detection by neutrons for a one minute detection period and one meter distance from the source, only 20 cm thick layer of lithium hydride would be needed \cite{Fetter1990}. To counter this, long detection period (\textgreater 10 mins) \cite{Fetter1990} would be needed. Avoid detection of the gamma rays is significantly more difficult. The most likely and effective material to block gamma rays is lead. In order to bring the required detection period from 1 meter away to a time greater than 5 mins, at least 100 kg of lead would be needed (see \autoref{fig:shield_time}).

@@

If short detection time periods are needed, then it is unlikely that passive signals could be effectively utilized. If longer detection time are able to be done, then gamma detection of weapons grade plutonium and uranium with a depleted uranium tempar are possible with distances greater than or equal to 1 meter away. However, the requirement of the tamper material to be depleted uranium makes this method of detection highly not recommended, as or tempar materials such as tungsten are not significantly less likely to be used.

@@

\textbf{Nuclear Resonance Fluorescense}

@@

\begin{table}[H]
\centering
\begin{tabular}{|c|c|c|}
\hline
Parameter                                                                                                                & \ce{^{235}U}                                                                                               & \ce{^{239}Pu}                                                                                               \\ \hline
\begin{tabular}[c]{@{}c@{}}Average Prompt Neutron Yield (multiplicity)\footnotemark[4]\\ (prompt neutrons per fission)\end{tabular}   & \begin{tabular}[c]{@{}c@{}}2.43 (thermal) \\ 2.57 (fission spectrum) \\  4.6 (14 MeV)\end{tabular} & \begin{tabular}[c]{@{}c@{}}2.87 (thermal) \\  3.09 (fission spectrum) \\  4.9 (14 MeV)\end{tabular} \\ \hline
\begin{tabular}[c]{@{}c@{}}Average Prompt Neutron Energy\footnotemark[5]\\ (MeV)\end{tabular}                                         & \begin{tabular}[c]{@{}c@{}}1.935 (thermal) \\  2.03 (14 Mev)\cite{PhysRevC.69.034607}\end{tabular}                       & \begin{tabular}[c]{@{}c@{}}2.010 (thermal) \\  2.19 (14 Mev)\cite{PhysRevC.69.034607}\end{tabular}                        \\ \hline
\begin{tabular}[c]{@{}c@{}}Average Prompt Photon Yield\\ (prompt photons per fission)\end{tabular}                       & 6.60\(\pm\)0.2 (thermal)\cite{Valentine2001}                                                                            & 7.06\(\pm\)0.2 (thermal)\cite{Valentine2001}                                                                             \\ \hline
\begin{tabular}[c]{@{}c@{}}Average Prompt Photon Energy\footnotemark[6]\\ (MeV)\end{tabular}                                          & 0.97\(\pm\)0.04 (thermal)\cite{Valentine2001}                                                                           & 0.95\(\pm\)0.04 (thermal)\cite{Valentine2001}                                                                            \\ \hline
\begin{tabular}[c]{@{}c@{}}Average Delayed Neutron Yield (multiplicity)\footnotemark[4]\\ (delayed neutrons per fission)\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.0158 (thermal) \\  0.0165 (1.45 MeV)\end{tabular}                     & \begin{tabular}[c]{@{}c@{}}0.0061 (thermal) \\  0.0063 (1.45 MeV)\end{tabular}                      \\ \hline
Average Delayed Neutron Energy (MeV)                                           & 0.43                                                                                               & 0.4                                                                                                 \\ \hline
\begin{tabular}[c]{@{}c@{}}Average Delayed Photon Yield\footnotemark[7]\\ (delayed photons per fission)\end{tabular}                  & \begin{tabular}[c]{@{}c@{}}0.613 (short period) \\  3.31 (long period)\end{tabular}                & \begin{tabular}[c]{@{}c@{}}0.608 (short period) \\  3.26 (long period)\end{tabular}                 \\ \hline
Average Delayed Photon Energy (MeV)\footnotemark[8]                                         & 0.96                                                                                               & 0.98                                                                                                \\ \hline
\end{tabular}
\caption[Signatures from Active Neutron Interrogation \cite{Chichester2009}]{Signatures from Active Neutron Interrogation \cite{Chichester2009} \footnotemark[3]}
\label{tab:neutron_signatures}
\end{table}

@@

The NRC defines LD50(30) to be the dose of radiation expected to cause death to 50 percent of an exposed population within 30 days (LD 50/30). Typically, the LD 50/30 is in the range from 400 to 450 rem (4 to 5 sieverts) received over a very short period \cite{U}.  From this metric, the dose of 400 rem will be selected for evaluation here, and would require a very large source emission of radiation - typically one much larger than the practical release and yield of a RDD - and it will not occur over a very short period.

@@

There are three approaches to disrupting ISIL'ss operations that will be particularly important in ensuring that ISIL does not acquire a nuclear device:

@@

In the current oil market, diminished profits from selling oil may make ISIL more dependent on external donations. As need for this arises, it is important to focus on these external sources. Saudi Arabia and UAE are becoming more effective at preventing funding of terrorist orgnaizations by their citizens, but Kuwait and Qatar remain permissive jurisdictions. More needs to be done by these states to enforce their newly passed laws against financing terrorism, and the U.S. can help them do this through State Department and other training. It is important to note that these Gulf states have an interest in stopping ISIL – ISIL has periodically targeted Qatar and its Ministry of Foreign Affairs. \subsubsection{???????? THIS PARAGRAPH NEEDS SOURCES   ?????????}

@@

Disabling ISIL oil revenue is more complicated than it might seem. ISIL has been focused on establishing a durable internal market for oil, ensuring they have a reliable source of fuel for its fleets and also establishing dependence from the local population on its cheap oil. Recent airstrikes have targeted the oil at its source, but might be better served by targeting means of transportation and sale. Syrian rebels have pointed to this as a major error in targeting, as it disrupts oil supply to power generators and local facilities, allowing ISIL to pass the blame on to the U.S. (who they call \enquote{the crusaders}) when the electricity system failes. One difficulty in targeting the means of transportation and sale is that oil is transported in tanker trucks and oil drums, and most of the transactions are in cash, making them hard to track. One policy priority should be providing large quantities of diesel fuel and oil for generators in opposition areas of northern Syria [1].
A final recommendation is to put more intelligence resources toward detecting ISIL fundraising efforts through modern communication networks (social media) \cite{Report2015}.

@@

The U.S. should request countries to proactively identify individuals and entities for inclusion in the UN Al Qaida Sanctions Committee list \cite{Report2015}.

@@

\subsubsection{???????? not sure the first paragraph of this quote is necessary, unless your saying that we should spend comprable amounts of money now. also, you can probably just intergrate the points of the second paragraph into your own paragraph, no real need to pull it out as a block quote  ?????????}

@@

ISIL could attempt to purchase or steal a nuclear weapon, and so the nations which maintain a stockpile of nuclear weapons must remain under scrutiny. Nations of the highest interest that have nuclear weapons are Pakistan, North Korea, India, Turkey, Israel, and Russia. Israel has not been confirmed to have nuclear weapons, but it is widely believed that it has nuclear capabilites. Turkey is a NATO nuclear weapon sharing state and has U.S. tactical nuclear weapons stationed within its borders. However, it is unlikely that ISIL would be able to steal a nuclear weapon from either of these states, as those in Turkey are under U.S. protection and Israel has proven adept at hiding their arsenal. Of greatest concern is the possibility that any of these nuclear-capable states, North Korea in particular, would willingly sell a physics package to ISIL.

@@

Most of these instances are very small amounts (grams) but there have been several involving one to three kilograms. No individual case has been found with enough material to make a nuclear bomb, but muiltiple thefts could allow ISIL to accumulate enough special nuclear material to create a nuclear weapon.

@@

\subsubsection{RDD}
 Some of the materials required to build an RDD are naturally occurring elements that are naturally radioactive. Other components are solely availabile to radiological work sites such as reactor facilities, isotope production, research and universities, weapons production facilities, and other commercial and industrial facilities.

The lands surrounding ISIL are vast, and currently a large majority are under militant occupation.  The probability that materials of this nature can move about the region is very high.  Paid courier services, as well as stolen armed and armoured transportation vehicles  may assist in this endeavour.

@@

Agency cooperation to manage and coordinate a crisis. This would include current actions to prepare to respond to a crisis, and might have lessons learned from Fukushima and Ebola in terms of interagency action. How do we utilize agencies and individuals without access to top secret information in the event of a loose nuke? There will be a lot of expertise that could help but intel community and white house won't want broad knowledge that ISIL may be getting close to nuclear device.

@@

The information volume received from the all intelligence sources is astronomic. To facilitate the handling of this volume of information, assets are assigned based on the prioritization of the information. These prioritizations are as low, medium, and highly important:

>>>>

Where appropriate, expected countermeasures, such as shielding of radioactive sources, are discussed within the context of discerning a weak signal in a strong background. In circumstances where it is deemed essential for personnel to monitor the shipment of regular commerce or travel, including but not limited to commercial or private traffic by land, sea, or air, portal detectors and active interrogation is considered. Such locations are qualified for observation which are considered high risk or of strategic value in fielding a conventional response to attack from ISIL on the U.S., allies, or any sovereign territory for which instability is a matter of U.S. National Security. It is assumed that ISIL will prioritize counter-value over counterforce targets, since counter-value targeting is more disruptive and more effective in a campaign of terror.

@@

    Throughout the many conflicts in Iraq in the remainder of the 2000s, AQI underwent many changes in leadership, ideology, and name. These changes, primarily the deaths of much of its upper leadership (including al-Zarqawi), were caused by a combination of military pressure from U.S. troops, multiple disagreements  with bin Laden's al-Qaeda, and backlash from other Sunnis in Iraq \cite{Zelin2014,Kahl2008}. As a result of this, AQI was reorganized as the Islamic State of Iraq (ISI) in 2007, with Abu Bakr al-Baghdadi as it newly appointed leader \cite{Zelin2014,Shadid2010}. Under the new leadership of al-Baghdadi, ISI broke a period of regional détente with al-Qaeda (following the death of Osama bin Laden in August 2011), driving a further wedge between the two groups by al-Baghdadi's refusal to pledge allegiance to the new leader of al-Qaeda, Ayman al-Zawahiri \cite{Al-Jawlani}. In April 2013, this division resulted in  al-Baghdadi announcing that ISI would be extending its reach into Syria, and changing its name to the Islamic State of Iraq and al-Sham, or the Islamic State of Iraq and the Levant \cite{Zelin2014,Al-Hussaini2013}. In June 2014, al-Baghdadi declared ISIL a global Caliphate, with himself as its Caliph \cite{Mortada2014,TheWeek2014}. To understand the significance of this declaration, let us consider the history and role of Caliphs and Caliphates in Islam.

@@

ISIL has many enemies, indeed many more than are graphically depicted in \autoref{fig:ISILRelationships}.  As of January 2015, the anti-ISIL coalition led by the U.S. included 62 nations and groups including the U.S., Canada, Iraq, Jordan, Bahrain, Saudi Arabia, UAE, France, Germany, United Kingdom, Australia, Belgium, Denmark, Italy, Czech Republic, Albania, Netherlands, Estonia, Hungary, Turkey, and Lebanon, all of whom have carried out or supported military operations against ISIL \cite{Wordsworth2015}.  Additionally, although not allied with the U.S. coalition, Syria, Iran, and various Kurdish, Shia, and Sunni militias in Syria and Iraq have opposed ISIL \cite{Mooney2014}.  Despite this broad coalition, there are many unlikely allies in the list which present potential seams for ISIL exploitation  through propaganda or military action.  For example, strikes against ISIL in Syria benefit the Assad regime while U.S. trained rebels could turn on and attack the Assad regime \cite{Shinkman2015}.  Officially, the U.S. supports the removal of the Assad regime albeit without direct military intervention on part of the U.S. \cite{Shinkman2015}.  Similarly, U.S. air strikes and intelligence have supported Iran and Iraqi Sunni and Shia militias (including several that once fought U.S. forces) in their campaigns against ISIL \cite{Chulov2014}.  While these uneasy alliances have placed pressure on ISIL, it is unclear if they can be sustained over the timeframe that would be necessary to defeat ISIL and prevent their acquisition of a nuclear weapon. Sustaining these alliances would be especially difficult if ISIL were to turn its formidable propaganda machine to breaking some of these fragile links.

@@

The minimum amount of special nuclear material needed to produce a fission weapon is called the critical mass. This value can depend on many different features including isotope, geometry of the material, micro-structure of the material, and device design. Given the many variables in play it would be infeasible to include the critical mass of every possible weapons configuration here. However, to give a general idea, \autoref{tab:critical_mass} presents the critical mass of two different designs for several types of weapons. Additionally the yields of fission devices can vary based on many factors including, type of material, device design, amount, efficiency, and many other factors. Yields are typically on the order of kilotons \cite{Moody2014}.

@@

The most important step in the process, from a proliferation standpoint, involves enriching the uranium in \ce{^{235}U}. Natural uranium is typically 0.72\% \ce{^{235}U} \cite{Benedict1981}.  If uranium is destined for a power plant, it is typically enriched to 3-5\% \ce{^{235}U}. If on the other hand it is going to be used in a weapon, the uranium must be enriched to greater than approximately 90\% \ce{^{235}U} \cite{Moody2014}.

@@

It should be noted that Syria may have had a reactor for the production of plutonium in the past. However, that site was bombed and destroyed by an Israeli air strike in 2007 \cite{WorldNuclearAssociation2015a}. It is not impossible that ISIL could obtain the plans for that design. That being said, construction on a nuclear power plant typically takes at least 4 years \cite{NuclearEnergyInstitute2015}.  Furthermore, nuclear reactors are incredibly difficult to hide, as seen by Israel's 2007 air raid against Syria \cite{WorldNuclearAssociation2015a}.  Thus, it seems unlikely that ISIL would be able to complete such a reactor. Therefore, there is a very low probability that ISIL could covertly produce it's own plutonium, and thus little need to monitor for reprocessing signals from ISIL territory.

@@

From the U.S. perspective, 9/11 serves as a case study to understand the shortcomings of IC, subsequent recommended changes, and possible shortcomings facing the IC with respect to ISIL. The 9/11 Report found that there are cultural and organizational factors that hinder strategic communication between U.S intelligence agencies (See \autoref{app:partners} for a discussion on strategic communication on the international level, which also applies at the inter-agency level).  Prior to 9/11, U.S intelligence agencies operated in a culture dominated by \enquote{need-to-know}, which implied that organizations were given discretion in sharing data \cite{McConnell2008}. In the context of a dynamic ISIL threat, this means that agencies often have static analyses formed from an incomplete picture and lacking a competing analysis derived from the same data. The divisions between individual agencies render them incapable of accessing, prioritizing, translating, analyzing, and relaying the information for the policymakers to respond to in a timely manner. As a result, information flow is hindered by each agency's networks and databases lacking protocols to interact with other agencies' networks and databases.  Organizational inertia, standard operating procedures, and guidelines hinder the ability to change or implement new ideas. For example, the CIA's culture of risk aversion in the 1990s was detrimental to its intelligence capabilities and prevented them from properly assessing and taking action against the rising threat of terrorism \cite{Zegart2005}.

@@

Several international organizations and treaties are in place to facilitate communication flow and prevent the spread of nuclear weapons and technology (see \autoref{app:partners}).  However, communication in these organizations organization can be a challenge, and most of the current infrastructure is designed towards state actors, not the current ISIL threat.  Although not a specific alliance against ISIL or nuclear terrorism, the five eyes agreement (see \autoref{app:intel}) can provide valuable intelligence on possible nuclear activities that ISIL may engage. This group is stable, efficient, and more nimble in intelligence collaboration than the existing international non-proliferation structure.

@@

The number of countries in the coalition is growing with Obama's campaign, but the bar for inclusion is fairly low. While many countries have committed military capabilities to combating ISIL, some countries join the coalition by merely declaring their intent to contribute. All these member countries don't share the same understanding of whom the real enemy is - ISIL or the Syrian regime led by President Bashar al-Assad. For example, Turkey has urged the coalition to topple the Assad regime as a condition to provide its air bases to carry out attacks against ISIL.

@@

Iran and U.S. shared a common enemy factor in terms of ISIL. As the nuclear talks between the two countries are being hashed out, both countries have committed on fighting ISIL together. This can be a good opportunity for U.S. to combat ISIL more effectively as U.S. cannot disregard Iran's influence and role in Iraq's security landscape and their fight against ISIL.
Meanwhile, collaborating with Israel is considered a gray area and not clear-cut as it seems. Israel has been known to voice their support of ISIL, keeping in line with ISIL's common enemy with Israel i.e. Iran. As U.S. and Israel have been a strong ally in the Middle East region, this fact presented a conundrum to the U.S..

@@

If nuclear material is intercepted before detonation, an analysis of the enrichment and contaminants may be achieved at the National Labs (Oak Ridge, Los Alamos). On the scale of weeks to months, it is possible to discern the age of the material post fabrication by analyzing the micro-structure, activity, and isotopic composition. Immediately following detonation, the timescale of analysis relies on how quickly radiation monitoring equipment may be delivered to the blast site and how quickly samples are gathered and shipped back to one of 16 radiological facilities equipped to perform nuclear forensics \cite{1446400}. Airplanes capable of sampling large volumes of the dust and gas (Xenon) in the plume generated by the detonation must be performed within 1-2 days of the detonation and may indicate the composition, efficiency, and sophistication of the device. Analysis of materials taken from the ground near the detonation point will take 1-2 weeks to analyze and will confirm and refine the findings of atmospheric monitoring. The identification of the origin of the weapon will depend largely on what is presently known about the yield, sophistication, and efficiency of weapons in global stockpiles. What is exactly known of the capabilities of weapons designed outside of the United States is classified \cite{Glasstone_1964}. Given quick access to the nuclear blast site and a somewhat thorough understanding of the nuclear capabilities of other sovereign powers, the technical analysis of the origin and design of a nuclear blast would be available within 3-10 days, depending on the availability of data. Ongoing research is aimed at developing LIBS technologies which may reduce the nuclear forensics timeline with rapid (\textless 15 hours) debris dissolutions techniques, or techniques which do not require the dissolution of collected debris \cite{Condron}.

@@

While the other weapon models have a fair distance of detection for a short time, the data does not take into account, additional shielding. To prevent detection by neutrons for a one minute detection period and one meter distance from the source, only 20 cm thick layer of lithium hydride would be needed \cite{Fetter1990}. To counter this, long detection period (\textgreater 10 minutes) \cite{Fetter1990} would be needed. Avoid detection of the gamma rays is significantly more difficult. The most likely and effective material to block gamma rays is lead. In order to bring the required detection period from 1 meter away to a time greater than 5 minutes, at least 100 kg of lead would be needed (see \autoref{fig:shield_time}).

@@

If short detection time periods are needed, then it is unlikely that passive signals could be effectively utilized. If longer detection time are able to be done, then gamma detection of weapons grade plutonium and uranium with a depleted uranium tamper are possible with distances greater than or equal to 1 meter away. However, the requirement of the tamper material to be depleted uranium makes this method of detection highly not recommended, as or tamper materials such as tungsten are not significantly less likely to be used.

@@

\textbf{Nuclear Resonance Fluorescence}

@@

\begin{table}[H]
\centering
\begin{tabular}{|c|c|c|}
\hline
Parameter                                                                                                                & \ce{^{235}U}                                                                                               & \ce{^{239}Pu}                                                                                               \\ \hline
\begin{tabular}[c]{@{}c@{}}Average Prompt Neutron Yield (multiplicity)\footnotemark[4]\\ (prompt neutrons per fission)\end{tabular}   & \begin{tabular}[c]{@{}c@{}}2.43 (thermal) \\ 2.57 (fission spectrum) \\  4.6 (14 MeV)\end{tabular} & \begin{tabular}[c]{@{}c@{}}2.87 (thermal) \\  3.09 (fission spectrum) \\  4.9 (14 MeV)\end{tabular} \\ \hline
\begin{tabular}[c]{@{}c@{}}Average Prompt Neutron Energy\footnotemark[5]\\ (MeV)\end{tabular}                                         & \begin{tabular}[c]{@{}c@{}}1.935 (thermal) \\  2.03 (14 MeV)\cite{PhysRevC.69.034607}\end{tabular}                       & \begin{tabular}[c]{@{}c@{}}2.010 (thermal) \\  2.19 (14 MeV)\cite{PhysRevC.69.034607}\end{tabular}                        \\ \hline
\begin{tabular}[c]{@{}c@{}}Average Prompt Photon Yield\\ (prompt photons per fission)\end{tabular}                       & 6.60\(\pm\)0.2 (thermal)\cite{Valentine2001}                                                                            & 7.06\(\pm\)0.2 (thermal)\cite{Valentine2001}                                                                             \\ \hline
\begin{tabular}[c]{@{}c@{}}Average Prompt Photon Energy\footnotemark[6]\\ (MeV)\end{tabular}                                          & 0.97\(\pm\)0.04 (thermal)\cite{Valentine2001}                                                                           & 0.95\(\pm\)0.04 (thermal)\cite{Valentine2001}                                                                            \\ \hline
\begin{tabular}[c]{@{}c@{}}Average Delayed Neutron Yield (multiplicity)\footnotemark[4]\\ (delayed neutrons per fission)\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.0158 (thermal) \\  0.0165 (1.45 MeV)\end{tabular}                     & \begin{tabular}[c]{@{}c@{}}0.0061 (thermal) \\  0.0063 (1.45 MeV)\end{tabular}                      \\ \hline
Average Delayed Neutron Energy (MeV)                                           & 0.43                                                                                               & 0.4                                                                                                 \\ \hline
\begin{tabular}[c]{@{}c@{}}Average Delayed Photon Yield\footnotemark[7]\\ (delayed photons per fission)\end{tabular}                  & \begin{tabular}[c]{@{}c@{}}0.613 (short period) \\  3.31 (long period)\end{tabular}                & \begin{tabular}[c]{@{}c@{}}0.608 (short period) \\  3.26 (long period)\end{tabular}                 \\ \hline
Average Delayed Photon Energy (MeV)\footnotemark[8]                                         & 0.96                                                                                               & 0.98                                                                                                \\ \hline
\end{tabular}
\caption[Signatures from Active Neutron Interrogation \cite{Chichester2009}]{Signatures from Active Neutron Interrogation \cite{Chichester2009} \footnotemark[3]}
\label{tab:neutron_signatures}
\end{table}

@@

The NRC defines LD50(30) to be the dose of radiation expected to cause death to 50 percent of an exposed population within 30 days (LD 50/30). Typically, the LD 50/30 is in the range from 400 to 450 rem (4 to 5 Sv) received over a very short period \cite{U}.  From this metric, the dose of 400 rem will be selected for evaluation here, and would require a very large source emission of radiation - typically one much larger than the practical release and yield of a RDD - and it will not occur over a very short period.

@@

There are three approaches to disrupting ISIL's operations that will be particularly important in ensuring that ISIL does not acquire a nuclear device:

@@

In the current oil market, diminished profits from selling oil may make ISIL more dependent on external donations. As need for this arises, it is important to focus on these external sources. Saudi Arabia and UAE are becoming more effective at preventing funding of terrorist organizations by their citizens, but Kuwait and Qatar remain permissive jurisdictions. More needs to be done by these states to enforce their newly passed laws against financing terrorism, and the U.S. can help them do this through State Department and other training. It is important to note that these Gulf states have an interest in stopping ISIL – ISIL has periodically targeted Qatar and its Ministry of Foreign Affairs. \subsubsection{???????? THIS PARAGRAPH NEEDS SOURCES   ?????????}

@@

Disabling ISIL oil revenue is more complicated than it might seem. ISIL has been focused on establishing a durable internal market for oil, ensuring they have a reliable source of fuel for its fleets and also establishing dependence from the local population on its cheap oil. Recent airstrikes have targeted the oil at its source, but might be better served by targeting means of transportation and sale. Syrian rebels have pointed to this as a major error in targeting, as it disrupts oil supply to power generators and local facilities, allowing ISIL to pass the blame on to the U.S. (who they call \enquote{the crusaders}) when the electricity system fails. One difficulty in targeting the means of transportation and sale is that oil is transported in tanker trucks and oil drums, and most of the transactions are in cash, making them hard to track. One policy priority should be providing large quantities of diesel fuel and oil for generators in opposition areas of northern Syria [1].
A final recommendation is to put more intelligence resources toward detecting ISIL fundraising efforts through modern communication networks (social media) \cite{Report2015}.

@@

The U.S. should request countries to proactively identify individuals and entities for inclusion in the UN al-Qaeda Sanctions Committee list \cite{Report2015}.

@@

\subsubsection{???????? not sure the first paragraph of this quote is necessary, unless your saying that we should spend comparable amounts of money now. also, you can probably just integrate the points of the second paragraph into your own paragraph, no real need to pull it out as a block quote  ?????????}

@@

ISIL could attempt to purchase or steal a nuclear weapon, and so the nations which maintain a stockpile of nuclear weapons must remain under scrutiny. Nations of the highest interest that have nuclear weapons are Pakistan, North Korea, India, Turkey, Israel, and Russia. Israel has not been confirmed to have nuclear weapons, but it is widely believed that it has nuclear capabilities. Turkey is a NATO nuclear weapon sharing state and has U.S. tactical nuclear weapons stationed within its borders. However, it is unlikely that ISIL would be able to steal a nuclear weapon from either of these states, as those in Turkey are under U.S. protection and Israel has proven adept at hiding their arsenal. Of greatest concern is the possibility that any of these nuclear-capable states, North Korea in particular, would willingly sell a physics package to ISIL.

@@

Most of these instances are very small amounts (grams) but there have been several involving one to three kilograms. No individual case has been found with enough material to make a nuclear bomb, but multiple thefts could allow ISIL to accumulate enough special nuclear material to create a nuclear weapon.

@@

\subsubsection{RDD}
 Some of the materials required to build an RDD are naturally occurring elements that are naturally radioactive. Other components are solely available to radiological work sites such as reactor facilities, isotope production, research and universities, weapons production facilities, and other commercial and industrial facilities.

The lands surrounding ISIL are vast, and currently a large majority are under militant occupation.  The probability that materials of this nature can move about the region is very high.  Paid courier services, as well as stolen armed and armored transportation vehicles  may assist in this endeavor.

@@

Agency cooperation to manage and coordinate a crisis. This would include current actions to prepare to respond to a crisis, and might have lessons learned from Fukushima and Ebola in terms of inter-agency action. How do we utilize agencies and individuals without access to top secret information in the event of a loose nuke? There will be a lot of expertise that could help but intelligence community and white house won't want broad knowledge that ISIL may be getting close to nuclear device.

@@

The information volume received from the all intelligence sources is astronomic. To facilitate the handling of this volume of information, assets are assigned based on the prioritization of the information. These are assigned as low, medium, and highly important:
